\def\duedate{11/29/22}
\def\HWnum{4}
\input{../preamble.tex}
\usepackage[]{slashed}

\begin{document}
    
\prob{1}{
In the toy model introduced in lecture ($A$, $B$, and $C$ only)
}

a) Calculate the total cross section for $A \bar{A} \rightarrow A \bar{A}$ for the case $m_{A} = 0$, $m_{C} \gg E_{A}$.
Remember there are two diagrams for this process.

The tree level diagrams are shown in the following image, with momenta labeled:

For diagram 1, we have
\begin{eqnarray}
    \label{eq:amp1}
    \mathcal{M}_{1} = i(-ig)\frac{i}{s - m_{C}^2}(-ig) 
,\end{eqnarray}
where $s = (p_1 + p_2)^2$,
and for diagram 2 we have
\begin{eqnarray}
    \label{eq:amp2}
    \mathcal{M}_{2} = i(-ig)\frac{i}{t - m_{C}^2}(-ig) 
,\end{eqnarray}
where $t = (p_1 - p_3)^2$.
The total amplitude is given as
\begin{eqnarray}
    \label{eq:squared-amp}
    \mathcal{M} = \frac{g^{2}}{s - m_{C}^2} + \frac{g^{2}}{t - m_{C}^2}
.\end{eqnarray}
We can simplify the Mandelstam variables ($s$ and $t$) as 
\begin{eqnarray}
    \label{eq:mand-vars}
    \begin{aligned}    
        s &= p_1^2 + p_2^2 + 2 p_1 \cdot p_2 = \phantom{-} 2 p_1 \cdot p_2 = 4 E_{A}^2 \\
        t &= p_1^2 + p_3^2 - 2 p_1 \cdot p_3 = -2 p_1 \cdot p_3 = -2 E_{A}^2 (1 + \cos{\theta})
    .\end{aligned}
\end{eqnarray}
Note that $E_{A} = E_{\bar{A}} = | \mathbf{p} |$ since $m_{A} = m_{\bar{A}} = 0$.
The center of mass differential cross section is then
\begin{align}
    \label{eq:dxsec-CM}
    \dv{\sigma}{\Omega} &= \frac{1}{64 \pi^2 (E_{A} + E_{\bar{A}})^2} \frac{| \mathbf{p}' |}{| \mathbf{p} |} | \mathcal{M} |^2 \\
                        &= \frac{g^{4}}{64 \pi^2 (4 E_{A}^2)} \Big[ \frac{1}{s - m_{C}^2} + \frac{1}{t - m_{C}^2} \Big]^2 \\
                        &= \frac{g^{4}}{64 \pi ^2 E_{A}^2 m_{C}^{4}} \Big[ 1 + \Big( \frac{E_{A}}{m_{C}} \Big)^2 \sin^2{(\theta/2)} \Big]
,\end{align}
where $E_{A}$ and $E_{\bar{A}}$ are the initial state energies of the corresponding particles and $\mathbf{p}$ and $\mathbf{p}'$ are the initial and final state 3-momenta in the CM frame.
Note that the term in square brackets was expanded in a series of powers of $E_{A}/m_{C}$ and higher order terms were thrown away.
Integrating over the solid angle $\Omega$, we find the total cross section to be
\begin{eqnarray}
    \label{eq:tot-xsec-AAbar}
    \eqbox{
        \sigma = \frac{g^{4}}{16 \pi E_{A}^2 m_{C}^{4}} \Bigg[ 1 + \frac{1}{2} \Big( \frac{E_{A}}{2 m_{C}} \Big)^2 \Bigg]
}
.\end{eqnarray}


b) Calculate the total cross section for $A A \rightarrow A A$ for the case $m_{A} = 0$, $m_{C} \gg E_{A}$.
Remember that there are two diagrams for this process and that the integral should be taken only over half the solid angle or the $S$ factor should be used to avoid double counting.
What is the ratio of the cross sections for parts (a) and (b)?

The two possible diagrams are shown below:

As in part (a) we use our toy Feynman rules to find that
\begin{align}
    \label{eq:amps-b}
    \mathcal{M}_{1} &= \frac{g^2}{(p_1 - p_3)^2 - m_{C}^2} \\
    \mathcal{M}_{2} &= \frac{g^2}{(p_3 - p_2)^2 - m_{C}^2}
,\end{align}
which gives the total amplitude as
\begin{eqnarray}
    \label{eq:amp-b}
    \mathcal{M} = g^2 \Big[ \frac{1}{t - m_{C}^2} + \frac{1}{u - m_{C}^2} \Big]
\end{eqnarray}
and differential cross section as
\begin{eqnarray}
    \label{eq:dxsec-b}
    \dv{\sigma}{\Omega} = \frac{g^{4}}{128 \pi^2 E_{A}^2 m_{C}^{4}} \Bigg[ 1 - \Big( \frac{2E_{A}}{m_{C}} \Big)^2 \Bigg]
.\end{eqnarray}
We then obtain the total cross section by integrating over the solid angle:
\begin{eqnarray}
    \label{eq:xsec-b}
    \eqbox{
        \sigma = \frac{g^{4}}{32 \pi E_{A}^2 m_{C}^{4}} \Bigg[ 1 - \Big( \frac{2 E_{A}}{m_{C}} \Big)^2 \Bigg]
    } 
.\end{eqnarray}



c) Calculate the differential cross sections for $A \bar{A} \rightarrow A \bar{A}$ and $A A \rightarrow A A$ for the case $m_{A} = m_{C} = 0$.

Using our results for the amplitudes from parts (a) and (b) and inserting $m_{C} = 0$ we have
\begin{eqnarray}
    \label{eq:xsec1}
    \eqbox{
    \dv{\sigma(A \bar{A} \rightarrow A \bar{A})}{\Omega} = \frac{g^{4}}{64 \pi^2 (4 E_{A}^2)} \Big( \frac{1}{s} + \frac{1}{t} \Big)^2 = \Big( \frac{g^2 \tan^2{(\theta/2)}}{64 \pi E_{A}^3} \Big)^2
}
\end{eqnarray}
and
\begin{eqnarray}
    \label{eq:xsec2}
    \eqbox{
    \dv{\sigma(AA \rightarrow AA)}{\Omega} = \frac{g^{4}}{128 \pi^2 (4 E_{A}^2)} \Big( \frac{1}{t} + \frac{1}{u} \Big)^2 = \frac{1}{2} \Big( \frac{g^{2}}{16 \pi E_{A}^3 \sin^2{\theta}} \Big) ^2
}
.\end{eqnarray}


d) What is the ratio of the two cross sections for part (c) at very small angles?
What is the ratio for scattering at $90^{\circ}$ in the center of mass?
At $180^{\circ}$?

The ratio of the two cross sections is as follows:
\begin{eqnarray}
    \label{eq:ratio}
    R = \frac{\dd{\sigma(A \bar{A} \rightarrow A \bar{A})}}{\dd{\sigma(AA \rightarrow AA)}} = \frac{1}{8} \sin^{4}{\theta}\tan^{4}{(\theta/2)}
.\end{eqnarray}
At small angles, the ratio is
\begin{eqnarray}
    \label{eq:ratio-small-angle}
    R(\theta) = \frac{\theta^{8}}{128}
,\end{eqnarray}
which is incredibly small.

Furthermore,
\begin{eqnarray}
    \label{eq:ratio-90}
    R(90^{\circ}) = \frac{1}{8} 
,\end{eqnarray}
and
\begin{eqnarray}
    \label{eq:ratio-180}
    R(180^{\circ}) = \infty
.\end{eqnarray}



\prob{2}{
The Higgs boson has been discovered, cannot decay into two real $W$ or $Z$ pairs because its mass is too low.
However, consider another HIggs boson with a sufficiently high mass and spin 0.
Its coupling will be proportional to the Higgs mass divided by the $W$ mass, i.e. the vertex factor will be proportional to $-i g M_{H}/M_{W}$, where $g$ is a dimensionless number.
From dimensional analysis, how do you expect the Higgs decay width to vary with the unknown Higgs mass in this very high mass region?
}

\prob{3}{
Show that $S^{\dagger}S \ne 1$ but $S^{\dagger}\gamma^{0}S = \gamma^{0}$.
}

Recall that 
\begin{eqnarray}
    \label{eq:S-mat}
    S = \begin{pmatrix}
        a_{+} & a_{-}\sigma_1 \\
        a_{-}\sigma_1 & a_{+}
    \end{pmatrix}   
,\end{eqnarray}
where $a_{\pm} = \pm \sqrt{\frac{1}{2}(\gamma \pm 1)}$ and $\gamma = 1/\sqrt{1 - v^2}$ is the Lorentz boost factor (in natural units).
Thus,
\begin{eqnarray}
    \label{eq:SdaggerS}
    \begin{aligned}    
        S^{\dagger} S &= \begin{pmatrix}
        a_{+} & a_{-} \sigma_1 \\
        a_{-} \sigma_1 & a_{+}
    \end{pmatrix}
    \begin{pmatrix}
        a_{+} & a_{-} \sigma_1 \\
        a_{-} \sigma_1 & a_{+}
    \end{pmatrix}
    = 
    \begin{pmatrix}
        a_{+}^2 + a_{-}^2 \sigma_1^2 & 2 a_{+}a_{-} \sigma_1 \\
        2 a_{+}a_{-} \sigma_1 & a_{-}^2 \sigma_1^2 + a_{+}^2
    \end{pmatrix} 
    \\
                      &=
    \begin{pmatrix}
        \gamma & -\sqrt{\gamma^2 - 1} \sigma_1 \\
        -\sqrt{\gamma^2 - 1}\sigma_1 & \gamma
    \end{pmatrix}
    =
    \gamma
    \begin{pmatrix}
        1 & -v \sigma_1 \\
        -v \sigma_1 & 1
    \end{pmatrix} 
    \ne 1
    \end{aligned}
.\end{eqnarray}

Checking the second identity, we have
\begin{eqnarray}
    \label{eq:Sdagger-gamma0-S}
    S^{\dagger} \gamma^{0} S = 
    \begin{pmatrix}
        a_{+} & a_{-} \sigma_1 \\
        a_{-} \sigma_1 & a_{+}
    \end{pmatrix}
    \begin{pmatrix}
        1 & 0 \\
        0 & 1
    \end{pmatrix}
    \begin{pmatrix}
        a_{+} & a_{-} \sigma_1 \\
        a_{-} \sigma_1 & a_{+}
    \end{pmatrix}
    =
    \begin{pmatrix}
    a_{+}^2 - a_{-}^2\sigma_1^2 & 0 \\
    0 & -a_{+}^2 + a_{-}^2\sigma_1^2
    \end{pmatrix} 
    = \gamma^{0}
.\end{eqnarray}



\prob{4}{
Using the plane-wave solutions of the Dirac equation, verify the completeness relation 
\begin{eqnarray}
    \label{eq:completeness-spinors}
    \sum_{s=1,2} u^{(s)} \overline{u}^{(s)} = \slashed{p} + m
,\end{eqnarray}
where $\slashed{p} = \gamma^{\mu}p_{\mu}$ and natural units are used.
}

The spinors
\begin{eqnarray}
    \label{eq:particle-spinors}
    u^{(1)} = \sqrt{E + m}
    \begin{pmatrix}
        1 \\ 0 \\ \frac{p_{z}}{E + m} \\ \frac{p_{x} + ip_{y}}{E + m}
    \end{pmatrix} 
    \quad
    u^{(2)} = 
    \sqrt{E + m}
    \begin{pmatrix}
    0 \\ 1 \\ \frac{p_{x} - ip_{y}}{E + m} \\ - \frac{p_{z}}{E + m}
    \end{pmatrix} 
.\end{eqnarray}



\prob{5}{
In class we calculated the differential cross section for $e^{+}e^{-} \rightarrow \mu^{+}\mu^{-}$ and obtained 
\begin{eqnarray}
    \label{eq:e+e-mu+mu-_xsec}
    \dv{\sigma}{\Omega} = \frac{\alpha^2}{4s}(1 + \cos^2{\theta})
,\end{eqnarray}
where $s$ is the total energy squared in the center of mass, $\theta$ is the angle between the $e^{-}$ and $\mu^{-}$ in the center of mass, and where we ignored the electron and muon masses.
Redo the calculuation for $e^{+}e^{-} \rightarrow \tau^{+}\tau^{-}$ for an energy range in which the electron mass can be ignored, but no the $\tau$ mass .
In particular, what is the factor by which the \textit{total} cross section is suppressed by the $\tau$ mass for the case in which the colliding electron and positron each have an energy of \SI{2}{\GeV}.
}






\end{document}
